---
title: "Coding Challenge 5"
author: "Theresa Quintana"
date: "2025-03-20"
output:
  md_document:
      variant: gfm
  pdf_document:
  html_document:
    toc: true
    toc_float: true
  word_document:
---
You can find the corresponding documents to this assignment on my GitHub: 
[Theresa's Coding Challenge 5 Github Access](https://github.com/taq-poly/CodingChallenge5.git)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

#Question 1: Download two .csv files from Canvas called DiversityData.csv and Metadata.csv, and read them into R using relative file paths.
```{r}
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(ggrepel)
library(ggprism)
library(knitr)

diversity <- read.csv("DiversityData.csv", na.strings = "na")
meta <- read.csv("Metadata.csv", na.strings = "na")
```

#Question 2: Join the two dataframes together by the common column ‘Code’. Name the resulting dataframe alpha.
```{r}
alpha <- left_join(diversity, meta, by = "Code")
alpha
```

#Question 3: Calculate Pielou’s evenness index: Pielou’s evenness is an ecological parameter calculated by the Shannon diversity index (column Shannon) divided by the log of the richness column. 
```{r}
#a. Using mutate, create a new column to calculate Pielou’s evenness index. 
#b. Name the resulting dataframe alpha_even.
alpha_even <- mutate(alpha, PielouIndex = shannon/log(richness))
alpha_even
```

#Question 4: Using tidyverse language of functions and the pipe, use the summarise function and tell me the mean and standard error evenness grouped by crop over time.
```{r}
#a.	Start with the alpha_even dataframe
#b.	Group the data: group the data by Crop and Time_Point.
#c.	Summarize the data: Calculate the mean, count, standard deviation, and standard error for the even variable within each group.
#d.	Name the resulting dataframe alpha_average

alpha_average <- alpha_even %>% 
  group_by(Crop, Time_Point) %>% #group by Crop & Time_Point to later calculate summary stats by group
  summarise(mean.even = mean(PielouIndex), #calculate the mean, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(PielouIndex)) %>%
  mutate(std.err = sd.dev/sqrt(n)) 
alpha_average
```

#Question 5: Calculate the difference between the soybean column, the soil column, and the difference between the cotton column and the soil column
```{r}
#a.	Start with the alpha_average dataframe
#b.	Select relevant columns: select the columns Time_Point, Crop, and mean.even.
#c.	Reshape the data: Use the pivot_wider function to transform the data from long to wide format, creating new columns for each Crop with values from mean.even.
#d.	Calculate differences: Create new columns named diff.cotton.even and diff.soybean.even by calculating the difference between Soil and Cotton, and Soil and Soybean, respectively.
#e.	Name the resulting dataframe alpha_average2

alpha_average2 <- alpha_average %>%
  select(Time_Point, Crop, mean.even) %>%
  pivot_wider(names_from = Crop, values_from = mean.even) %>%
  mutate(diff.cotton.even = Soil - Cotton) %>%
  mutate(diff.soybean.even = Soil - Soybean)
alpha_average2
  
```

#Question 6: Connecting it to plots
```{r}
#a.	Start with the alpha_average2 dataframe
#b.	Select relevant columns: select the columns Time_Point, diff.cotton.even, and diff.soybean.even.
#c.	Reshape the data: Use the pivot_longer function to transform the data from wide to long format, creating a new column named diff that contains the values from diff.cotton.even and diff.soybean.even.
#d.	Create the plot: Use ggplot and geom_line() with ‘Time_Point’ on the x-axis, the column ‘values’ on the y-axis, and different colors for each ‘diff’ category. The column named ‘values’ come from the pivot_longer. The resulting plot should look like the one to the right. 

Plot <- alpha_average2 %>%
  select(Time_Point, diff.cotton.even, diff.soybean.even) %>%
  pivot_longer(c(diff.cotton.even, diff.soybean.even), names_to = "diff") %>%
  ggplot(aes(x = Time_Point, y = value, color = diff)) + # add the ggplot
  geom_line() +
  xlab("Time (hrs)") +
  ylab("Differences from Soil in Pielou's Evenness") +
  theme_prism()
Plot

```

#Question 7.	Commit and push a gfm .md file to GitHub inside a directory called Coding Challenge 5. Provide me a link to your github written as a clickable link in your .pdf or .docx


